---
title: "Analysis of Happiness Score Dataset"
author: "Prasham Bhuta <br /> github.com/Prashambhuta"
date: "16/03/2020"
mainfont: Calibri Light
output:
  pdf_document: default
  html_document:
    df_print: paged
sansfont: Calibri Light
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dataset

Dataset ranks 155 countries, from the period: 2015 to 2017, by their happiness through 6 indicators.

* economic production
* family
* health
* freedom
* trust (government corruption)
* generosity

The final indicator is dystopia residual. Dystopia is a made up country which ranks lowest in all of the above 6 criteria. The score generated in such a way is used to compare dystopia residual with happiness score.

* low dystopia residual = low happiness score
* high dystopia residual = high happiness score

The dataset is cleaned and provided courtesy [Kian](https://twitter.com/CoderKianWee)

## My Exercise

* To generate plots of indicators vs happiness score.
* To calculate Paerson's correlation coefficient and analyse the dependency of happiness score across all indicators.

## Let's Get Started

**Importing necessary library**
```{r, echo=TRUE, message=FALSE}
library(readr)
library(dplyr)
```

**Generating data from csv**
```{r}
data <- read.csv("data_sets/World_Happiness_2015_2017.csv", stringsAsFactors = FALSE, header = T)
```

**Analysing the data**
```{r}
str(data)
```

**Cleaning column headers**
```{r}
colnames(data) <-(c("Country", "Happiness.Rank", "Happiness.Score", "Economy", "Family", "Health", "Freedom","Trust", "Generosity", "Dystopia.Residual", "Year"))
str(data)
```


## Correlation between Economy and Happiness Score
```{r}
corr_econ <- cor(data %>% select(Economy, Happiness.Score), method= "pearson")
cat("corr_econ (R):", corr_econ[2], "\nR^2:", corr_econ[2]^2)
data_econ <- data %>% select(Happiness.Score, Economy)
plot(data_econ, main="Economy vs Happiness Score", col="darkgreen", type="p", pch=20, cex=1.2)
abline(lm(data$Economy ~ data$Happiness.Score), col="green", lwd=3)
```

The correlation coefficient (R) is `0.785` and the square (R<sup>2</sup>) is `0.616`.

Happiness score is strongly related to the economy of the country. High economy means

* higher living standards, 
* less economical disparity,
* monetary satisfaction across all line of work.

## Family vs Happiness Score
```{r}
data_fam <- data %>% select(Happiness.Score, Family)
corr_fam = cor(data_fam, method="pearson")[2]
cat("R:", corr_fam, "\nR^2:", corr_fam^2)
plot(data_fam, col="darkgreen", pch=20, cex=1.3)
abline(lm(data_fam$Family ~ data_fam$Happiness.Score), col="green", lwd=3)
```
R is `0.63` while R<sup>2</sup> is `0.40`, this shows that Happiness Score is strongly dependent on Family. Happy members of family contribute to increase in Family indicator and Happiness Score.

## Health vs Happiness Score
```{r}
data_health <- data %>% select(Happiness.Score, Health)
corr_health <- cor(data_health, method = "pearson")[2]
cat("R:", corr_health, "\nR^2:", corr_health^2)
plot(data_health, col="darkgreen", pch=20, cex=1.3)
abline(lm(data_health$Health ~ data_health$Happiness.Score), col="green", lwd=3)
```

The R<sup>2</sup> coefficient is `0.55`, this indicates that health has positive relation to Happiness. People with higher life expectancy are happier compared to countries with lower life expectancy.

## Freedom vs Happiness Score
```{r}
data_freedom <- data %>% select(Happiness.Score, Freedom)
corr_freedom <- cor(data_freedom, method = "pearson")[2]
cat("R:", corr_freedom, "\nR^2:", corr_freedom^2)
plot(data_freedom, col="darkgreen", pch=20, cex=1.3)
abline(lm(data_freedom$Freedom ~ data$Happiness.Score), col="orange", lwd=3)
```

The R<sup>2</sup> coefficient for Freedom vs Happiness score is `0.31`. Compared to Economy, Family and Health; the coefficient is lower. However as per the data, happiness has strong dependence on Freedom. Freedom includes freedom to act upon a wish, such as changing careers, choosing partners, starting new business etc. People which such liberty are believed to be happier.

## Trust vs Happiness Score
```{r}
data_trust <- data %>% select(Happiness.Score, Trust)
corr_trust <- cor(data_trust, method = "pearson")[2]
cat("R:", corr_trust, "\nR^2:", corr_trust^2)
plot(data_trust, col="darkgreen", pch=20, cex=1.3)
abline(lm(data_trust$Trust ~ data$Happiness.Score), col="red", lwd=3)
```

The R<sup>2</sup> coefficient for Trust on government is `0.07`. This indicates that the role of a government in the happiness of an human being is low. As long as government boosts the economy, and provides support & health to the families, people are happy.

## Generosity vs Happiness Score
```{r}
data_generosity <- data %>% select(Happiness.Score, Generosity)
corr_generosity <- cor(data_generosity, method = "pearson")[2]
cat("R:", corr_generosity, "\nR^2:", corr_generosity^2)
plot(data_generosity, col="darkgreen", pch=20, cex=1.3)
abline(lm(data_generosity$Generosity ~ data$Happiness.Score), col="red", lwd=3)
```

The R<sup>2</sup> coefficient for Generosity (`0.04`) is the lowest among all indicators. Generosity or helpfulness of other people has no to minimum say in the overall happiness of the society.

## Dystopia Residual vs Happiness Score
```{r}
data_dys_res <- data %>% select(Happiness.Score, Dystopia.Residual)
corr_dys_res <- cor(data_dys_res, method = "pearson")[2]
cat("R:", corr_dys_res, "\nR^2:", corr_dys_res^2)
plot(data_dys_res, col="darkgreen", pch=20, cex=1.3)
abline(lm(data_dys_res$Dystopia.Residual ~ data$Happiness.Score), col="orange", lwd=3)
```

The R<sup>2</sup> coefficient for Dystopia Residual is `0.24`. Dystopia Residual means how much the said country is doing better than the worst country i.e. `Dystopia`. It is strong indicator and shows that countries with averages higher than the median are more happier than country whose indicator score lie close to that of Dystopia.

## Conclusion
**From the data we can conclude which indicators are major influence on happiness, and which indicators are negeligible. The following table depicts the dependency:**

|Level of Dependency |Indicator 1| Indicator 2| Indicator 3|
|---|-----|-----|-----|
|Strongly Dependent | Economy (`0.62`) | Health (`0.55`) | Family (`0.40`) |
|Mildly Dependent | Freedom (`0.31`) | Dystopia Residual (`0.24`)| - |
|Weakly Dependent | Trust [Government] (`0.08`) | Generosity (`0.05`) | - |

[Check out the Jupyter Notebook for the same](https://colab.research.google.com/drive/1bn7oXAwqrmHsoJo4fLrb5e8XORcXSy3K)